RCloud: Integrating Exploratory Visualization, Analysis and Deployment

### 2014-03-26 discussion.tex
working

What are the decisions and ideas that are central to this paper?

discussion about binding times, running vs.caching.
transparent vs. explicit operations done by programmers

% limitations
Previous studies have pointed out difficulties in achieving the flexibility,
scalability and maintainability expected of production software in experimental
code written by data analysts.
Although these properties cannot be enforced by any programming environment,
suitable tools can help well-motivated analysts and programmers to create
robust applications with much less effort.

@@

Had to back off only one language sooner than we expected.

Shared artifacts, artifact histories, 
Discussion
View sharing - bookmarking, everything available as a URL
Content export - want to avoid, create ways of working where this is not necessary.
Social-psychological incentives - starring
Voting and ranking - starring

Group management, size, diversity.  Lacking support,but clearly desirable.

Has to play well in the ecosystem of the other tools and worked out as planned but there is a temptation to absorb

@@

>>>>

\subsection{Design Considerations}

Experience with deploying RCloud in a community of data analysts
has provided some insight into how its capabilities were received
that may be informative to researchers and designers of future systems
supporting visual analytics.
Returning to the design considerations outlined by Heer and Agrawala,
our experience underscores the value of several of their proposals,
and suggests areas for further exploration.

{\bf Shared artifacts and artifact histories} are a central feature.
The prototype aims at making it convenient to collaborate on code
and experiments with data. It was a success to observe that analysts
readily shared ideas through RCloud. We observed them creating experiments
in workbooks and sharing URLs to discuss data analysis and programming
problems, instead of sending screenshots in Powerpoint.

{\bf View sharing, bookmarking.} Likewise, everything is named
and accessed as a URL.

{\bf Discussion.} Annotation is supported in our human interface,
and was also a central goal for our prototype.
An interesting question is to decide to what extent the users of
production notebooks also use tools like annotation etc. without coding
or without being exposed to the analyst's view.\todo{Is this a gap in
our design?}

{\bf Content export} is not a capability we worked to support,
but the R environment already has many packages for it.
Our goal is to devise ways of working where explicit export is
less necessary. However it is desirable to integrate results
through web protocols such as JSON. We did assume that RCloud
notebooks need to play well in the ecosystem of other tools
and this worked out as planned, though it also leads to a temptation
to absorb features almost without limit. (Should RCloud support
templates for personal web pages? Should it incorporate emacs
on the web, and an email client?)

{\bf Social-psychological incentives} and {\bf voting and ranking}
are supported through starring. This mechanism was employed often
by our test audience. An obvious next step would be to enhance
recommendations using relationships discovered by static and
dynamic code analysis. This should be done both within and across
collections of scripts (the latter being similar to work in Vistrails
that enhanced recommendations by clustering multiple workflows).
It may be valuable to data analysts to know which packages are often
used together, or with a given type of data. It is probably helpful
to know which data sources and packages often go together.

{\bf Group management, size and diversity.} This area lacks sufficient
support in RCloud, but is clearly desirable. We rely on conventional
administrative and social processes, outside our platform. Automation
of this capability in a large organization could rely on social media
services

\subsection{Limitations}

Because we developed the ideas and the prototype together,
we could not forsee some of the requirements that emerged
after our audience started working with the prototype.

It was unexpected that we had to had to back off so soon on
the assumption that programming is being done in only one
language (R). In retrospect, it should have been obvious that
many data science problems are already being solved with
cross-language approaches.

A clear disdavantage is that some important aspects of the environment in
RCloud are more difficult to manage than in conventional systems. Not only
do we lack an explicit separation of the development and deployment
environments, but every version of every prototype is shared by default.
While this encourages collaboration on work in progress, errors such as
misconfiguration, mismatches between versions of packages,or ordinary
programming errors are also easily shared and could affect production
websites unintentionally. We take the position that this cannot entirely
be avoided, but the power of sharing is worth this price (just as that
has proven true with distributed systems in general). Still, this 
problem needs more attention. 
Versioning occurs in several places: in scripts, in the R environment
such as the installed R libraries and packages, and in the external
environent such as the operating system and even protocols spoken by
remote services. At one end of the scale, we have full control over the
versioning of scripts via git, though there is still a need for conventions
for example to name stable or working versions of scripts or versions known
to work together. The R environment is under the control of its package
manager. It is at least possible for RCloud to access information about
package versions and configurations, but support for checking compatibility
and maintaining multiple versions in the same environment is not strong.
At the other end of the scale, there is not much reason to expect version
control for the external environment. Most programmers rely on clean living
and a conservative approach to upgrades.

On top of this, RCloud allows analysts to fork experiments to try
new ideas quickly. But we have not done much so far to address the
probem of having a lot of bits and pieces of code and data lying
around, though at least we have made it easier to get at them and
to apply algorithms and metadata to organize them.

@@

@@

Discussion about binding times, running vs.caching.

Transparent vs. explicit operations done by programmers
